The 2014 SMT Competition

### 2015-08-01 2014/report/SMTCOMP.tex
More edits per review comments

In planning the 2014 competition, the organizers' overall goal was to encourage breadth
in the capability of SMT solvers. SMT-COMP 2014 benefited from the evaluation that was performed in 2013 and the experience of previous competitions. As a result we established the following emphases:
\begin{itemize}
\item In 2012 the competition was narrowed to a smaller number of more significant logics. In response to feedback, in 2014 we reverted to the practice of evaluating solvers in all available divisions.
\item We ran the competition with all available, eligible benchmarks. To be eligible, a benchmark must be well-formed and have a known answer. We were able to use all benchmarks because we had nearly exclusive use of the StarExec computational cluster during the time we were executing SMT-COMP. This point is discussed further in \S\ref{sec:benchmark-selection}
\item In 2012 some experimental tracks were held: parallel performance, unsat cores, and proof generation. The participation in those tracks was light. Since in 2014 we also had to migrate to using the new StarExec cluster to execute the competition, we held only a main track and an application (incremental) track in the 2014 competition. The organizers still appreciate the value of measuring the performance of solvers on new features such as model generation, unsat core determination, and proof generation, and recommend that they be reinstated in some future edition of the competition.
\item A significant change in 2014 was to increase the timeout limit for a solver processing a given benchmark from 25 minutes (in 2012) to 40 minutes (see \S\ref{sec:procedure} for details on the competition parameters and \S\ref{sec:timeouts} for a discussion on their effects).
\item A significant result of the 2013 Evaluation was that the result of a competition was highly sensitive to the selection of benchmarks: different random selections of benchmarks resulted in different winning orders in a large fraction of samples. Hence, an aim for 2014 was to use as large a benchmark set as possible in the competition to minimize this effect.

@@

The SMT Competition is a competition among SMT solvers on a set of benchmark logic problems. Each benchmark problem is a combination of definitions and logical assertions expressed with respect to an underlying logical theory and, perhaps, some constraints on the kinds of expressions in that theory. For example,  the logic of linear arithmetic includes the multiplication operation, but only allows multiplication by constants. Each problem is a set of closed formulas over a set of constant or function symbols, possibly including quantification; a solution to the problem is an assignment of each constant and function evaluation to values in a way that satisfies all the problem's assertions. That is, the task is to find a \emph{satisfying assignment} for the benchmark problem or to demonstrate that there is no such assignment. Since the presence of quantified expressions introduces incompleteness, solvers may also produce a potential solution that may be marked as possibly incomplete.

@@

However, it is important to note that some tools used in the competition support only a subset of the SMT-LIB command language, leaving out some commands that are essentially syntactic sugar, and are defined as optional in the standard to simplify the burden on competing solvers. This subset is only informally stated and has been stable since 2010. Moreover, the benchmarks only use this subset. This informality did not cause problems in the competition, but it would be preferable to define this subset explicitly. 
%% [DD] Strictly speaking, these commands are optional and a SMT-LIB
%% compliant solver needs not implement them. Which is why I commented
%% out the following sentence 
%%
%% Requiring just a subset for competition works against the goal of
%% producing SMT-LIB-compliant solvers for users.  Fortunately, most
%% mature solvers have implemented nearly all commands of SMT-LIB.
Optional features such as incrementality, proof production, and determining unsat cores are evaluated in specialized tracks, separate from the main competition. The presence of such tracks in the competition has varied from year to year.

@@

SMT-LIB currently defines 34 logics, shown in Table \ref{Table:benchmarks}. The rightmost column shows the number of benchmarks for that logic in the SMT-LIB collection. Two considerations may make a benchmark ineligible for a competition. First, the benchmark may not have a known result. For newly submitted benchmarks
we made an attempt to determine the correct result of the benchmark; the SMT-LIB coordinators require that two
different solvers solve the benchmark and report the same result. However, for benchmarks with unknown results already in the collection we did not have time to determine their results. We did do this analysis after the competition was over (see~\S\ref{sec:post}).

@@

@@

Each benchmark is assigned a \emph{difficulty rating}; in 2014, we used the time taken to solve the benchmark by the best performing solver in the 2013 SMT Evaluation. These values were publicly available prior to the competition. The difficulty ratings are used to divide the benchmarks into 5 quintiles. The rules describe a procedure for randomly selecting $N$ out of the eligible benchmarks for a division, with the intent of selecting roughly equal numbers, if they are available, from each of the quintiles.  The seed for the random number generator used for selection is obtained by summing a number supplied by each solver team and the integer portion of the New York stock exchange index at its opening on the day the competition begins. \davidc{Check the rules}

@@

The competition registration includes information about each competing solver. In addition, some solver groups provided summaries of their solvers and their recent technical advances.
 Note that although one person is listed as the ``submitter,'' there is generally a team of contributors behind each tool. Some teams submitted more than one tool. The 2014 participants were the following:
\begin{itemize}
\item 4Simp -- submitted by Trevor Hansen (U. Melbourne)
\item AbzizPortfolio -- two versions -- submitted by Mohammed Adbul Aziz (U. Cairo). This solver is atypical in that it is a portfolio solver: based on automated learning over benchmark characteristics, it chooses among other solvers to apply to the problem at hand.
\item AProVe~\cite{AProVE2014} -- submitted by Carsten Fuhs (University College London)
\item Boolector~\cite{Boolector2015} -- three versions: Boolector (default),  Boolector-d (dual propagation), Boolector-j (justification) -- submitted by Armin Biere, Aina Niemetz, Mathias Preiner (Johnnes Kepler University)
\item CVC3 v. 2.4.3~\cite{BT07} -- submitted by Morgan Deters (New York University)
\item CVC4 v. 1.4~\cite{BCD+11} -- submitted by the ACSys Group (New York University)
\item Kleaver -- 2 versions -- submitted by Hristina Palikareva, Cristian Cadar (Imperial College)
\item OpenSMT2 -- submitted by Antti Hyv\"arinen (U.~Lugano)
\item raSAT~\cite{DBLP:conf/smt/KhanhVO14} -- submitted by Xuan-Tung Vu (Japan Advanced Insitute of Science and Technology)
\item SMTInterpol~\cite{DBLP:conf/spin/ChristHN12,DBLP:conf/spin/2012} -- submitted by Jochen Hoenicke, J\"urgen Christ (U.~Freiburg)
\item SONOLAR~\cite{Peleska:2011:ATC:1986308.1986333} -- submitted by Florian Lapschies (U.~Bremen)
\item STP-CryptoMiniSat4~\cite{DBLP:conf/cav/GaneshD07,DBLP:conf/cav/2007} -- submitted by Mate Soos (Security Research Lab.), based on previous work by Trevor Hansen (U.~Melbourne) and Vijay Ganesh (Massachusetts Institute of Technology)
\item veriT~\cite{veriT} -- submitted by David D\'{e}harbe (UFRN - Universidade Federal do Rio Grande do Norte), Pascal Fontaine (U.~Lorraine, INRIA - Institut national de recherche en informatique et en automatique)
\item Yices2~\cite{Dutertre:cav2014} -- submitted by Bruno Dutertre (SRI)
\end{itemize}

@@

During the testing period prior to the competition beginning, different solvers exhibited different behavior on some benchmarks. Investigation revealed that these differences were the result of different treatments of partial operations, divide by zero, in particular. The question is what should be the definition of divide by zero in SMT-LIB's logic of total functions. This question had been discussed at length over previous years, with proposals to clarify the SMT-LIB standard.
There was certainly not time to resolve this matter prior to the competition, not to mention correcting solvers to the agreed behavior.\footnote{The discussion this time around can be found in the archives of the \url{smtcomp-discussion@lists.sourceforge.net} mailing list, in the few days near 2014-06-16.}
Thus, for the competition, we decided to omit all benchmarks (6060 of them, 1348 of which were otherwise eligible) that were affected by divide-by-zero behavior. This question, however, remains open.

@@

SMT differs from the CASC competition~\cite{PSS02} in directly addressing sorted logics. SMT also focuses on fragments of first-order logic that are \emph{decidable}. For example, a subset of the benchmarks are problems in an integer-difference logic, for which there are specific decision procedures. The competition among tools is to create very efficient implementations of a breadth of decision procedures. Accordingly, SMT solvers have historically not handled logics with quantified expressions, though several solvers now do so, responding to strong encouragement from industrial users to include quantification in SMT solvers. This can, in general, make the satisfiability problem undecidable and the solvers necessarily rely on heuristics. Solvers may respond that the answer to a problem is unknown; that is, although there may be a candidate satisfying value assignment, the solver cannot assure that some combination of instantiations of quantified formulas might not invalidate the assignment.
Nevertheless, the inclusion of logics with quantification has been welcome by users. With SMT solvers moving to include quantification and CASC solvers moving to include types and arithmetic, we see a basis for fruitful future collaboration between these communities.

@@

\davidc{Update the reference for SLCOMP}
\davidd{Added a journla submission on SLCOMP 2014}

Besides the subject material, the most interesting difference among competitions is in scoring. (The scoring procedure for
SMT-COMP is described in detail in section \ref{sec:floc}.)
Competitions are currently mostly focused on raw solver capability---that is, the numbers and difficulty of problems that can be solved. Thus competition winners are primarily determined by the total number of problems solved correctly,
with the time taken to produce the solutions used only as a tie-breaker. This is the case for SMT-COMP, SAT, and SV-COMP,
for example.

A more significant issue is how solver errors are taken into account in the metrics. A solver error indicates an unsoundness in the tool, which is highly undesirable; however, highly penalizing errors may discourage new entrants whose tools are not yet mature. Different competitions vary on this aspect. SMT-COMP has historically emphasized soundness; a
single error in a division caused the solver to be scored lower than any other solver with no errors, even if the other solvers solved very few problems. For the SAT competition, an error outright disqualifies a tool in the category it occurs\footnote{See \url{http://satcompetition.org/2014/rules.shtml}.}. In contrast, in the SV-competition~\cite{svcomp-2015}, successful runs account positively towards the final score, while errors count negatively (with an increased weight).

@@

>>>>

In planning the 2014 competition, the organizers' overall goal was to encourage breadth
in the capability of SMT solvers. SMT-COMP 2014 benefited from the evaluation that was performed in 2013 and the experience of previous competitions. As a result we established the following emphases. Note that, as described in later sections, the competition is
divided into a number of divisions, each focuses on a given logic and each with its own set of benchmark problems and solvers.
\begin{itemize}
\item In 2012 the competition was narrowed to a smaller number of more significant logics. In response to feedback, in 2014 we reverted to the practice of evaluating solvers in all available divisions.
\item A significant result of the 2013 Evaluation was that the result of a competition was highly sensitive to the selection of benchmarks: different random selections of benchmarks resulted in different winning orders in a large fraction of samples. Hence, an aim for 2014 was to use as large a benchmark set as possible in the competition to minimize this effect. We were able to run the competition with all eligible benchmarks. This point is discussed further in \S\ref{sec:benchmark-selection}
\item In 2012 some experimental tracks were held: parallel performance, unsat cores, and proof generation. The participation in those tracks was light. Since in 2014 we also had to migrate to using the new StarExec cluster to execute the competition, we held only a main track and an application (incremental) track in the 2014 competition. The organizers still appreciate the value of measuring the performance of solvers on new features such as model generation, unsat core determination, and proof generation, and recommend that they be reinstated in some future edition of the competition.
\item A significant change in 2014 was to increase the timeout limit for a solver processing a given benchmark from 25 minutes (in 2012) to 40 minutes (see \S\ref{sec:procedure} for details on the competition parameters and \S\ref{sec:timeouts} for a discussion on their effects).

@@

The SMT Competition is a competition among SMT solvers on a set of benchmark logic problems. Each benchmark problem is a combination of definitions and logical assertions expressed with respect to an underlying logical theory and, perhaps, some constraints on the kinds of expressions in that theory. For example,  the logic of linear arithmetic includes the multiplication operation, but only allows multiplication by constants. Each problem is a set of closed formulas over a set of constant or function symbols, possibly including quantification; a solution to the problem is an assignment of each constant and function evaluation to values in a way that satisfies all the problem's assertions. That is, the task is to find a \emph{satisfying assignment} for the benchmark problem or to determine that there is no such assignment. Since the presence of quantified expressions introduces incompleteness, solvers may also produce a potential solution that may be marked as possibly incomplete.

@@

%However, it is important to note that some tools used in the competition support only a subset of the SMT-LIB command language, leaving out some commands that are essentially syntactic sugar, and are defined as optional in the standard to simplify the burden on competing solvers. This subset is only informally stated and has been stable since 2010. Moreover, the benchmarks only use this subset. This informality did not cause problems in the competition, but it would be preferable to define this subset explicitly. 
%% [DD] Strictly speaking, these commands are optional and a SMT-LIB
%% compliant solver needs not implement them. Which is why I commented
%% out the following sentence 
%%
%% Requiring just a subset for competition works against the goal of
%% producing SMT-LIB-compliant solvers for users.  Fortunately, most
%% mature solvers have implemented nearly all commands of SMT-LIB.
Optional features such as incrementality, proof production, and determining unsat cores are evaluated in specialized tracks, separate from the main competition. The presence of such tracks in the competition has varied from year to year.

@@

@@

The 34 logics are shown in Table \ref{Table:benchmarks}. The rightmost column shows the number of benchmarks for that logic in the SMT-LIB collection. Two considerations may make a benchmark ineligible for a competition. First, the benchmark may not have a known result. For newly submitted benchmarks
we made an attempt to determine the correct result of the benchmark; the SMT-LIB coordinators require that two
different solvers solve the benchmark and report the same result. However, for benchmarks with unknown results already in the collection we did not have time to determine their results. We did do this analysis after the competition was over (see~\S\ref{sec:post}).

@@

Each benchmark is assigned a \emph{difficulty rating}; in 2014, we used the time taken to solve the benchmark by the best performing solver in the 2013 SMT Evaluation. These values were publicly available prior to the competition. The difficulty ratings are used to divide the benchmarks into 5 quintiles. The rules describe a procedure for randomly selecting $N$ out of the eligible benchmarks for a division, with the intent of selecting roughly equal numbers, if they are available, from each of the quintiles.  The seed for the random number generator used for selection is obtained by summing a number supplied by each solver team and the integer portion of the New York Stock Exchange Composite Index at its opening on the day the competition begins.

@@

The competition registration requires participants to submit information about each competing solver. In addition, some solver groups provided summaries of their solvers and their recent technical advances.
 Note that although one person is listed as the ``submitter,'' there is generally a team of contributors behind each tool. Some teams submitted more than one tool. The 2014 participants were the following:
\begin{itemize}
\item 4Simp -- submitted by Trevor Hansen (U. Melbourne)
\item AbzizPortfolio -- two versions -- submitted by Mohammed Adbul Aziz (U. Cairo). This solver is atypical in that it is a portfolio solver: based on automated learning over benchmark characteristics, it chooses among other solvers to apply to the problem at hand.
\item AProVe~\cite{AProVE2014} -- submitted by Carsten Fuhs (University College London)
\item Boolector~\cite{Boolector2015} -- three versions: Boolector (default),  Boolector-d (dual propagation), Boolector-j (justification) -- submitted by Armin Biere, Aina Niemetz, Mathias Preiner (Johnnes Kepler University)
\item CVC3 v. 2.4.3~\cite{BT07} -- submitted by Morgan Deters (New York University)
\item CVC4 v. 1.4~\cite{BCD+11} -- submitted by the ACSys Group (New York University)
\item Kleaver -- 2 versions -- submitted by Hristina Palikareva, Cristian Cadar (Imperial College)
\item OpenSMT2 -- submitted by Antti Hyv\"arinen (U.~Lugano)
\item raSAT~\cite{DBLP:conf/smt/KhanhVO14} -- submitted by Xuan-Tung Vu (Japan Advanced Insitute of Science and Technology)
\item SMTInterpol~\cite{DBLP:conf/spin/ChristHN12,DBLP:conf/spin/2012} -- submitted by Jochen Hoenicke, J\"urgen Christ (U.~Freiburg)
\item SONOLAR~\cite{Peleska:2011:ATC:1986308.1986333} -- submitted by Florian Lapschies (U.~Bremen)
\item STP-CryptoMiniSat4~\cite{DBLP:conf/cav/GaneshD07,DBLP:conf/cav/2007} -- submitted by Mate Soos (Security Research Lab.), based on previous work by Trevor Hansen (U.~Melbourne) and Vijay Ganesh (Massachusetts Institute of Technology)
\item veriT~\cite{veriT} -- submitted by David D\'{e}harbe (UFRN - Universidade Federal do Rio Grande do Norte), Pascal Fontaine (U.~Lorraine, INRIA - Institut national de recherche en informatique et en automatique)
\item Yices2~\cite{Dutertre:cav2014} -- submitted by Bruno Dutertre (SRI)
\end{itemize}

@@

\paragraph{Undefined Behavior.} During the testing period prior to the competition beginning, different solvers exhibited different behavior on some benchmarks. Investigation revealed that these differences were the result of different treatments of partial operations, divide by zero, in particular. The question is what should be the definition of divide by zero in SMT-LIB's logic of total functions. This question had been discussed at length over previous years, with proposals to clarify the SMT-LIB standard.
There was certainly not time to resolve this matter prior to the competition, not to mention correcting solvers to the agreed behavior.\footnote{The discussion this time around can be found in the archives of the \url{smtcomp-discussion@lists.sourceforge.net} mailing list, in the few days near 2014-06-16.}
Thus, for the competition, we decided to omit all benchmarks (6060 of them, 1348 of which were otherwise eligible) that were affected by divide-by-zero behavior. This question, however, remains open.

@@

\paragraph{CASC} SMT differs from the CASC competition~\cite{PSS02} in directly addressing sorted logics. SMT also focuses on fragments of first-order logic that are \emph{decidable}. For example, a subset of the benchmarks are problems in an integer-difference logic, for which there are specific decision procedures. The competition among tools is to create very efficient implementations of a breadth of decision procedures. Accordingly, SMT solvers have historically not handled logics with quantified expressions, though several solvers now do so, responding to strong encouragement from industrial users to include quantification in SMT solvers; there are benchmarks and competition divisions corresponding to logics with quantification. Including quantified expressions can, in general, make the satisfiability problem undecidable and the solvers necessarily rely on heuristics. Solvers may respond that the answer to a problem is unknown; that is, although there may be a candidate satisfying value assignment, the solver cannot assure that some combination of instantiations of quantified formulas might not invalidate the assignment.
Nevertheless, the inclusion of logics with quantification has been welcome by users. With SMT solvers moving to include quantification and CASC solvers moving to include types and arithmetic, we see a basis for fruitful future collaboration and perhaps merging between these communities.

@@

\davidc{Update the reference for SLCOMP}
\davidd{Added a journla submission on SLCOMP 2014}
\davidc{Add other solvers, SV-COMP?}

\paragraph{SL-COMP} Besides the subject material, the most interesting difference among competitions is in scoring. (The scoring procedure for
SMT-COMP is described in detail in section \ref{sec:floc}.)
Competitions are currently mostly focused on raw solver capability---that is, the numbers and difficulty of problems that can be solved. Thus competition winners are primarily determined by the total number of problems solved correctly,
with the time taken to produce the solutions used only as a tie-breaker. This is the case for SMT-COMP, SAT, and SV-COMP,
for example.

\paragraph{Scoring differences.} A more significant issue is how solver errors are taken into account in the metrics. A solver error indicates an unsoundness in the tool, which is highly undesirable; however, highly penalizing errors may discourage new entrants whose tools are not yet mature. Different competitions vary on this aspect. SMT-COMP has historically emphasized soundness; a
single error in a division caused the solver to be scored lower than any other solver with no errors, even if the other solvers solved very few problems. For the SAT competition, an error outright disqualifies a tool in the category it occurs\footnote{See \url{http://satcompetition.org/2014/rules.shtml}.}. In contrast, in the SV-competition~\cite{svcomp-2015}, successful runs account positively towards the final score, while errors count negatively (with an increased weight).

@@

\davidc{Revise the paragraph above per review comments}

\davidc{More assessment needed}
